Cultural Transmission Can Explain the Evolution of Cooperative Behavior

### 2020-07-10 ms/main.tex
Remove Yoav's comment on eq.20
Add more explainations

For simplicity, assume that $T_A = T_B = T$ so we will get
\begin{equation} \label{eq:equal_horizontal_transmission}
% check this, I think you are missing a wbar on the left side, ie it should be
% \bar{\hat{p}' - \hat{p}} =  \hat{p} ^2(-\alpha bvT + cv(1-T)) + \hat{p} (\alpha bvT -cv(1-T))
  \hat{p}' - \hat{p} =  \hat{p} ^2(-\alpha bvT + cv(1-T)) + \hat{p} (\alpha bvT -cv(1-T))
\end{equation}

We are looking for the equilibrium $\hat{p}^*$ which occurs when $\hat{p}' = \hat{p}$.
It is easy to verify that fixation of either phenotype is an equilibrium, that is,
\begin{equation} \label{eq:result2.1}
  \hat{p}^*_0 =  0, \, \hat{p}^*_1 = 1 .
\end{equation}
We wish to determine when fixation of cooperation is locally stable. % state the general condition on the derivative that implies local stability
To do that, we calculate the derivative of $f(\hat{p}) = \hat{p}'-\hat{p}$, 
\begin{equation} \label{eq:derivative_of_phattag-phat}
 f'(\hat{p}) = 2\cdot\hat{p} (-\alpha bvT + cv(1-T)) + (\alpha bvT -cv(1-T)).
\end{equation}
For $\hat{p} = \hat{p}^*_1=1$, the derivative is
$f'(1) = -\alpha bvT + cv(1-T)$, and therefore, fixation of cooperation is locally stable when
\begin{equation} \label{eq:oblique_and_vertic_result2}
  \frac{b}{c}>\frac{1-T}{\alpha T}
\end{equation}
Which is the same condition we get in corollary 1.1.

@@

We now study the general case where $T_A \neq T_B$. We have
\begin{equation} \label{eq:general_case_polynomial}% check, because I think wbar is missing
  f(\hat{p}) = \hat{p}'-\hat{p} = \beta_1p^3 + \beta_2 p^2 + \beta_3p + \beta_4
\end{equation}
where 
\begin{equation} \label{eq:polynomial_coefficients}
\begin{split}
  & \beta_1 = c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B) \\
  & \beta_2 = (b(1-\alpha v)-c(1-v)-1)(T_A-T_B)-\alpha bv T_A + cv(1-T_B) \\
  & \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) \\
  & \beta_4 = 0
\end{split}
\end{equation}
Since $f(\hat{p})$ is a cubic polynomial, three equilibria exist.
We already know that $\hat{p}^*_0 = 0 $ and $\hat{p}^*_1 = 1$ are equilibria.
The third equilibrium can be found by solving $f(\hat{p})/\big(\hat{p}(1-\hat{p})\big)=0$, and it is
\begin{equation} \label{eq:oblique_and_vertic_result}
  \hat{p}^*_2 =  \frac{\alpha bvT_A + cvT_B + T_A-T_B-cv}{c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B)}
\end{equation}

We would like to know when $\hat{p}^*_2$ is legitimate equilibrium, i.e. $0<\hat{p}^*_2<1$.  
To do that we find the derivative  of $f(\hat{p})$,
\begin{equation} \label{eq:derivative_of_phattag-phat2}
  f'(\hat{p}) = 3\beta_0 \hat{p}^2 + 2\beta_1\hat{p}^2 + \beta_3
 \end{equation}
We assume that $T_A>T_B$ and we get that $\beta_1<0$ since $c(1-v)-b(1-\alpha v)<0$ and $c<b$ $v<1$. % why?
In this case the cubic $f(\hat{p})$ goes to infinity when $\hat{p}$ goes to infinity. 
Therefore, if $\hat{p}^*_2$ is positive then $f'(0)$ must be negative % are you sure? I think it must be positive, because i think that the two extremes of a cubic must go in opposite directions
\begin{equation} \label{eq:derivative_of_phattag-phat_p=0}
  f'(0) = \beta_3
 \end{equation}
$\hat{p}^*_2>0$ when $\beta_3<0$.
In addition, $\hat{p}^*_0=0$ is stable when $|f'(0)|<1$ therefore, when $-1<\beta_3<1$.
We can show that
\begin{equation} \label{eq:beta3-condition}
  \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) > -cv(1-T_B) > -1
\end{equation}
Therefore, when $\hat{p}^*_2>0$ we have $-1<\beta_3<0$ and $\hat{p}^*_0=0$ is locally stable.

>>>>

For simplicity, assume that $T_A = T_B = T$ so we will get
\begin{equation} \label{eq:equal_horizontal_transmission}
  \hat{p}' - \hat{p} =  \hat{p} ^2(-\alpha bvT + cv(1-T)) + \hat{p} (\alpha bvT -cv(1-T))
\end{equation}

We are looking for the equilibrium $\hat{p}^*$ which occurs when $\hat{p}' = \hat{p}$.
It is easy to verify that fixation of either phenotype is an equilibrium, that is,
\begin{equation} \label{eq:result2.1}
  \hat{p}^*_0 =  0, \, \hat{p}^*_1 = 1 .
\end{equation}
We wish to determine when fixation of cooperation is locally stable. 
An equilibrium point $\hat{p}^*$ is locally stable when the derivative of $f(\hat{p}) = \hat{p}'-\hat{p}$ at $p^*$ is negative, $f'(\hat{p}^*)<0$. 
Calculating the derivative of $f(\hat{p})$ give us 
\begin{equation} \label{eq:derivative_of_phattag-phat}
 f'(\hat{p}) = 2\cdot\hat{p} (-\alpha bvT + cv(1-T)) + (\alpha bvT -cv(1-T)).
\end{equation}
For $\hat{p} = \hat{p}^*_1=1$, the derivative is
$f'(1) = -\alpha bvT + cv(1-T)$, and therefore, fixation of cooperation is locally stable when
\begin{equation} \label{eq:oblique_and_vertic_result2}
  \frac{b}{c}>\frac{1-T}{\alpha T}
\end{equation}
Which is the same condition we get in corollary 1.1.

@@

We now study the general case where $T_A \neq T_B$. We have
\begin{equation} \label{eq:general_case_polynomial}
  f(\hat{p}) = \hat{p}'-\hat{p} = \beta_1p^3 + \beta_2 p^2 + \beta_3p + \beta_4
\end{equation}
where 
\begin{equation} \label{eq:polynomial_coefficients}
\begin{split}
  & \beta_1 = [c(1-v) -b (1-\alpha v)] \cdot(T_A-T_B) \\
  & \beta_2 = (b(1-\alpha v)-c(1-v)-1)(T_A-T_B)-\alpha bv T_A + cv(1-T_B) \\
  & \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) \\
  & \beta_4 = 0
\end{split}
\end{equation}
Since $f(\hat{p})$ is a cubic polynomial, three equilibria exist.
We already know that $\hat{p}^*_0 = 0 $ and $\hat{p}^*_1 = 1$ are equilibria.
The third equilibrium can be found by solving $f(\hat{p})/\big(\hat{p}(1-\hat{p})\big)=0$, and it is
\begin{equation} \label{eq:oblique_and_vertic_result}
  \hat{p}^*_2 =  \frac{\alpha bvT_A + cvT_B + T_A-T_B-cv}{c(1-v)\cdot(T_A-T_B) -b (1-\alpha v) \cdot(T_A-T_B)}
\end{equation}

We would like to know when $\hat{p}^*_2$ is legitimate equilibrium, i.e. $0<\hat{p}^*_2<1$.  
To do that we find the derivative  of $f(\hat{p})$,
\begin{equation} \label{eq:derivative_of_phattag-phat2}
  f'(\hat{p}) = 3\beta_0 \hat{p}^2 + 2\beta_1\hat{p}^2 + \beta_3
 \end{equation}
We assume that $T_A>T_B$ and we look on $\beta_1$:
\begin{equation} \label{eq:beta1}
  \begin{split}
   \beta_1 & = [c(1-v) -b (1-\alpha v)] \cdot(T_A-T_B) \\
   &< [c(1-\alpha v) - b(1-\alpha v)]\cdot(T_A-T_B) \\
   &= (1-\alpha v)(c-b)(T_A-T_B)
  \end{split}
 \end{equation}
From eq.~\ref{eq:beta1} we get that $\beta_1<0$ since $c-b<0$ $1-\alpha v>0$ and we assumed that $T_A > T_B$.
In this case the cubic $f(\hat{p})$ goes to infinity when $\hat{p}$ goes to minus infinity ($-\infty$). 
Therefore, if $\hat{p}^*_2$ is positive then $f'(0)$ must be negative % are you sure? I think it must be positive, because i think that the two extremes of a cubic must go in opposite directions DC: Yest, but I've fixed in the line before, there was minus sign missing
\begin{equation} \label{eq:derivative_of_phattag-phat_p=0}
  f'(0) = \beta_3
 \end{equation}
$\hat{p}^*_2>0$ when $\beta_3<0$.
In addition, $\hat{p}^*_0=0$ is stable when $|f'(0)|<1$ therefore, when $-1<\beta_3<1$.
We can show that
\begin{equation} \label{eq:beta3-condition}
  \beta_3 = \alpha bvT_A - cv(1-T_B) + (T_A-T_B) > -cv(1-T_B) > -1
\end{equation}
Therefore, when $\hat{p}^*_2>0$ we have $-1<\beta_3<0$ and $\hat{p}^*_0=0$ is locally stable.
